Solutions to Integer Programming from Quantum Annealing

### 2020-05-11 paper/main.tex
remove one sentence about dephasing
Started to write abstract

\begin{abstract}
Abstract here...
\end{abstract}

@@

For general graphs, existing "classical" algorithms either find minimal solutions in exponential time $\sim O( 1.5^n)$ \cite{Fomin2009, vanRooij2009} or approximate solutions in polynomial time. For example, greedy algorithms locally optimize decisions about which nodes to add to the dominating set.
Thus one is guaranteed to find a dominating set but not necessarily a MDS.
We demonstrate an example greedy algorithm in figure Fig.~\ref{fig:mds-greedy}.
\begin{figure*}
	\centering
	\begin{tabular}{p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}}
	\includegraphics[width=0.20\textwidth]{./figures/greedy-1.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-2.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-3.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-4.pdf}\\
	\centering\bf{a} & \centering\bf{b} & \centering\bf{c} & \centering\bf{d}
	\end{tabular}
	\caption{
		Example of greedy MDS algorithm:
		The algorithm iteratively adds vertices to the dominating set (black nodes) such that newly added nodes are maximally unconnected (white nodes).
		Connected nodes are neighbors of the DS (turquoise) or already in the DS.
		In (a), one finds three unconnected nodes [(2), (4), (7)] which are adjacent to three unconnected nodes each.
		In the first step $a\to b$, randomly the node (4) is picked.
		In (b), one has two unconnected choices [(1), (2)] with each adjacent to another unconnected node.
		After randomly selecting (2) when going from $b \to c$, only one unconnected node is remaining (6).
		The found solution is a (minimal) dominating set but not a MDS.
		To find the MDS using this greedy algorithm, the first node choice must be either (2) or (5).
	}
	\label{fig:mds-greedy}
\end{figure*}

\subsubsection{Minimum Dominating Set}
The solution to the minimum independent set can be expressed as an integer optimization problem given by,
\begin{align}
f(x) = &\min(\sum_{\nu \in V} \psi^x_{\nu}),\\
\textrm{subject to} \quad & \psi^x_{\nu} + \sum_{\mu \in \mathit{N}(\nu)} \psi^x_{\mu} \geq 1,\\
& \psi^x_{\nu} \in \{0, 1\}
\end{align}
where the dimension of the dependent variable $x$ is the number of vertices $\overline{\overline{V}}$.
The problem minimizes the number of vertices in $D$ with a binary variable $\psi^x$ encoded by a single qubit, subject to the constraint that at least one vertex in $\mathit{N}(\nu)$ is in $D$. For each vertex in $V$ we introduce slack variables $s_{\nu} = T^s_{\nu \alpha} \psi^s_{\alpha}$ to encode the inequality constraint such that
\begin{align}
f(x) = &\min(\sum_{\nu\in V} \psi^x_{\nu}),\\
\textrm{subject to} \quad & \psi^x_{\nu} + J_{\nu \mu} \psi^x_{\mu}- T^s_{\nu \alpha} \psi^s_{\alpha}  - 1 = 0,\\
& \psi_{\nu} \in \{0, 1\},\\
& 0 \leq s_{\nu} \leq |\mathrm{N}(\nu)|\\
& s_{\nu} \in \mathbb{Z},
\end{align}
where the nearest-neighbor sum is expressed by $J$ (symmetric and zero diagonal), the adjacency matrix for $G$.
The algorithm uses $N_q = \overline{\overline{V}} + \sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits---$\overline{\overline{V}}$ qubits to encode the vertices and another $\sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits to embed the slack variables. Therefore, the embedding at worse scales with $\overline{\overline{V}} \log_2 |\overline{\overline{V}}|$ qubits for fully connected graphs.

@@

In quantum annealing, the ground state of the problem Hamiltonian is obtained from the evolution of a time-dependent Hamiltonian,
\begin{equation}
H(s) = A(s) H^{\textrm{init}} + B(s) H^{\textrm{problem}}, \label{eq:tdhamiltonian}
\end{equation}
where $H^\textrm{init}=\sum_i\sigma^x_i$ on the DWave, and $H^\textrm{problem}$ is given by Eq.~(\ref{eq:matrix_form}) for the MDS application. The variable $s\in [0, 1]$ is the normalized annealing time and is allowed to vary between 1$\mu s$ to 2$ms$ on the DWave. On DWave solvers, annealing offsets effectively advance or delay the annealing schedule of individual qubits. In Fig.~\ref{fig:anneal_schedule}, the default DWave annealing schedule is shown in addition to the effects of applying positive and negative offsets. Note that due to the way the offsets are implemented on the hardware level, systematic errors are introduced since the initial and final values of $A(s)$ and $B(s)$ will differ for qubits with different offsets. Fig.~\ref{fig:anneal_schedule} shows the maximum offset that we use for the study, which translates to approximately (due to non-linearity) a 5\% advancement (blue) or delay (red) in the schedule.

@@

To solve for time evolution dynamics of quantum annealing including thermal and the decoherence effects, we solve for the master equation in Lindblad form
\begin{align}
\partial_t \rho (t) =  \frac{-i}{\hbar} [H(t) , \rho(t)] + \sum_j (2L_j \rho(t) L_j^\dagger - \{ L^\dagger_j L_j, \rho(t) \}) ,
\end{align}
where $\rho (t)$ is the density matrix at time $t$. $H(t)$ is the time-dependent Hamiltonian 
\begin{align}
\label{eq:annealH}
 H_{anneal}(t)&  =  - \sum_i  A_i(t)\sigma^x_i & \notag \\ 
 +&  \sum_{i>j} \sqrt{B_{i}(t)B_{j}(t)} J_{ij} \sigma^z_i \sigma^z_j +\sum_i B_{i}(t) h_i \sigma^z_i  ,&
\end{align}
where $A_i(t)$ and $B_{i}(t)$ are site-dependent annealing schedule functions. The site dependency takes into account of the annealing offset. $[,]$ denotes the Lie bracket. $L_j$ are the Lindblad operators for decoherence. $\{, \}$ denotes the anti-commutator. In this study we only include a minimal model for decoherence, i.e., the amplitude damping for non-interacting qubits. The Lindblad operator is $L_j=\sqrt{\gamma} \sigma^{+}_j$ for $h_j>0$ and $L_j=\sqrt{\gamma} \sigma^{-}_j$ for $h_j<0$, where $\gamma = 1/T_1$ is the decoherence rate for relaxation time $T_1$. The amplitude damping model also includes the dephasing effect, where the dephasing time is fixed to $T_2=\frac{T_1}{2}$. For simplicity we assume all qubits having the same relaxation time. The initial condition is the Gibbs canonical ensemble
\begin{equation}
\rho (0) =  \frac{e^{-\beta H(0)}}{\mbox{Tr}[e^{-\beta H(0)}]} ,
\end{equation}
where $\beta$ is the inverse temperature.
The probability to get the ground state at measurement is
\begin{align}
P =  \mbox{Tr} \{  \rho (t) \pi_{\mbox{gnd}} \}  ,
\end{align}
where the projection operator onto the degenerated ground states subspace is defined as $\pi_{\mbox{gnd}}=\sum_{i\in G} |\mbox{gnd}_i\rangle \langle \mbox{gnd}_i| $. Here $\{ | \mbox{gnd}_i \rangle | i \in G \}$ forms an orthonormal basis for the degenerated ground states subspace, i.e., $\langle \mbox{gnd}_j | \mbox{gnd}_i \rangle = \delta_{ij}$.

>>>>

\begin{abstract}
While quantum computing proposes promising solutions to computational problems not accessible with classical approaches, due to current hardware constraints, most quantum algorithms are not yet capable of computing systems of practical relevance, and classical counterparts outperform them.
To practically benefit from quantum architecture, one has to identify problems and algorithms with favorable scaling and improve on corresponding limitations depending on available hardware.
For this reason, we developed an algorithm that solves integer linear programming problems (ILP), a classically NP-hard problem, on a quantum annealer (QA) and investigated problem and hardware-specific limitations.
This work presents the formalism of how to map ILP problems to the annealing architectures. It illustrates the effects of decorrelation and many-body localization, for the minimum dominating set problem, and compares annealing results against numerical simulations of the quantum architecture.
We find that the QA ILP algorithm outperforms random guessing but is limited to small problems and that annealing schedules can be adjusted to reduce the effects of decorrelation--which is qualitatively confirmed by simulations. 
%Conclusion: What do these findings mean to us?
[Rephrase] We need better hardware to challenge heuristic solvers but we identified how to optimize hardware.
%Perspectives: What should be done next?
[Rephrase] Better understand correlations and localizations and improve simulation to scale to larger problems.

\end{abstract}

@@

For general graphs, existing "classical" algorithms either find minimal solutions in exponential time $\sim O( 1.5^n)$ \cite{Fomin2009, vanRooij2009} or approximate solutions in polynomial time. For example, greedy algorithms locally optimize decisions about which nodes to add to the dominating set.
Thus one is guaranteed to find a dominating set but not necessarily a MDS.
We demonstrate an example greedy algorithm in figure Fig.~\ref{fig:mds-greedy}.
\begin{figure*}
	\centering
	\begin{tabular}{p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}}
	\includegraphics[width=0.20\textwidth]{./figures/greedy-1.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-2.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-3.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-4.pdf}\\
	\centering\bf{a} & \centering\bf{b} & \centering\bf{c} & \centering\bf{d}
	\end{tabular}
	\caption{
		Example of greedy MDS algorithm:
		The algorithm iteratively adds vertices to the dominating set (black nodes) such that newly added nodes are maximally unconnected (white nodes).
		Connected nodes are neighbors of the DS (turquoise) or already in the DS.
		In (a), one finds three maximally unconnected nodes [(2), (4), (7)] which are adjacent to three unconnected nodes each.
		In the first step $a\to b$, randomly the node (4) is picked.
		In (b), one has three unconnected choices [(1), (2)] with each adjacent to another unconnected node and one node [(6)] which is only adjacent to neighbors of the DS.
		After randomly selecting (2) when going from $b \to c$, only one unconnected node is remaining (6).
		The found solution is a (minimal) dominating set but not a MDS.
		To find the MDS using this greedy algorithm, the first node choice must be either (2) or (5).
	}
	\label{fig:mds-greedy}
\end{figure*}

\subsubsection{Minimum Dominating Set}
The solution to the minimum independent set can be expressed as an integer optimization problem given by,
\begin{align}
f(x) = &\min\left(\sum_{\nu \in V} \psi^x_{\nu}\right),\\
\textrm{subject to} \quad & \psi^x_{\nu} + \sum_{\mu \in \mathit{N}(\nu)} \psi^x_{\mu} \geq 1,\\
& \psi^x_{\nu} \in \{0, 1\}
\end{align}
where the dimension of the dependent variable $x$ is the number of vertices $\overline{\overline{V}}$.
The problem minimizes the number of vertices in $D$ with a binary variable $\psi^x$ encoded by a single qubit, subject to the constraint that at least one vertex in $\mathit{N}(\nu)$ is in $D$. For each vertex in $V$ we introduce slack variables $s_{\nu} = T^s_{\nu \alpha} \psi^s_{\alpha}$ to encode the inequality constraint such that
\begin{align}
f(x) = &\min(\sum_{\nu\in V} \psi^x_{\nu}),\\
\textrm{subject to} \quad & \psi^x_{\nu} + J_{\nu \mu} \psi^x_{\mu}- T^s_{\nu \alpha} \psi^s_{\alpha}  - 1 = 0,\\
& \psi_{\nu} \in \{0, 1\},\\
& 0 \leq s_{\nu} \leq |\mathrm{N}(\nu)|\\
& s_{\nu} \in \mathbb{Z},
\end{align}
where the nearest-neighbor sum is expressed by $J$ (symmetric and zero diagonal), the adjacency matrix for $G$.
The algorithm uses $N_q = \overline{\overline{V}} + \sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits---$\overline{\overline{V}}$ qubits to encode the vertices and another $\sum_{\nu \in V} \log_2 \mathit(N(\nu))$ qubits to embed the slack variables. Therefore, the embedding at worse scales with $\overline{\overline{V}} \log_2 |\overline{\overline{V}}|$ qubits for fully connected graphs.

@@

In quantum annealing, the ground state of the problem Hamiltonian is obtained from the evolution of a time-dependent Hamiltonian,
\begin{equation}
H(s) = A(s) H^{\textrm{init}} + B(s) H^{\textrm{problem}}, \label{eq:tdhamiltonian}
\end{equation}
where $H^\textrm{init}=\sum_i\sigma^x_i$ on the DWave, and $H^\textrm{problem}$ is given by Eq.~(\ref{eq:matrix_form}) for the MDS application. The variable $s\in [0, 1]$ is the normalized annealing time and is allowed to vary between 1$\mu s$ to 2$ms$ (CK: time unit mapping) on the DWave. On DWave solvers, annealing offsets effectively advance or delay the annealing schedule of individual qubits. In Fig.~\ref{fig:anneal_schedule}, the default DWave annealing schedule is shown in addition to the effects of applying positive and negative offsets. Note that due to the way the offsets are implemented on the hardware level, systematic errors are introduced since the initial and final values of $A(s)$ and $B(s)$ will differ for qubits with different offsets. Fig.~\ref{fig:anneal_schedule} shows the maximum offset that we use for the study, which translates to approximately (due to non-linearity) a 5\% advancement (blue) or delay (red) in the schedule.

@@

To solve for time evolution dynamics of quantum annealing including thermal and the decoherence effects, we solve for the master equation in Lindblad form
\begin{align}
\partial_t \rho (t) =  \frac{-i}{\hbar} [H(t) , \rho(t)] + \sum_j (2L_j \rho(t) L_j^\dagger - \{ L^\dagger_j L_j, \rho(t) \}) ,
\end{align}
where $\rho (t)$ is the density matrix at time $t$. $H(t)$ is the time-dependent Hamiltonian 
\begin{align}
\label{eq:annealH}
 H_{anneal}(t)&  =  - \sum_i  A_i(t)\sigma^x_i & \notag \\ 
 +&  \sum_{i>j} \sqrt{B_{i}(t)B_{j}(t)} J_{ij} \sigma^z_i \sigma^z_j +\sum_i B_{i}(t) h_i \sigma^z_i  ,&
\end{align}
where $A_i(t)$ and $B_{i}(t)$ are site-dependent annealing schedule functions. The site dependency takes into account of the annealing offset. $[,]$ denotes the Lie bracket. $L_j$ are the Lindblad operators for decoherence. $\{, \}$ denotes the anti-commutator. In this study we only include a minimal model for decoherence, i.e., the amplitude damping for non-interacting qubits. The Lindblad operator is $L_j=\sqrt{\gamma} \sigma^{+}_j$ for $h_j>0$ and $L_j=\sqrt{\gamma} \sigma^{-}_j$ for $h_j<0$, where $\gamma = 1/T_1$ is the decoherence rate for relaxation time $T_1$. For simplicity we assume all qubits having the same relaxation time. The initial condition is the Gibbs canonical ensemble
\begin{equation}
\rho (0) =  \frac{e^{-\beta H(0)}}{\mbox{Tr}[e^{-\beta H(0)}]} ,
\end{equation}
where $\beta$ is the inverse temperature.
The probability to get the ground state at measurement is
\begin{align}
P =  \mbox{Tr} \{  \rho (t) \pi_{\mbox{gnd}} \}  ,
\end{align}
where the projection operator onto the degenerated ground states subspace is defined as $\pi_{\mbox{gnd}}=\sum_{i\in G} |\mbox{gnd}_i\rangle \langle \mbox{gnd}_i| $. Here $\{ | \mbox{gnd}_i \rangle | i \in G \}$ forms an orthonormal basis for the degenerated ground states subspace, i.e., $\langle \mbox{gnd}_j | \mbox{gnd}_i \rangle = \delta_{ij}$.
